---
title: "Streamflow Analysis"
author: 
- "Pablo E. Guti&eacute;rrez-Fonseca"
- "`r format(Sys.time(), '%d %B, %Y')`"
layout: default
permalink:
output: html_document
---
<br/>
<br/>


#### **Step 1**
Load libraries.<br>
_Cargar las librerias que necesitas._
```{r, warning=FALSE, message=FALSE}
libraries <- c('hydroTSM','dataRetrieval','data.table','dplyr','ggplot2','lubridate','raster','rgdal','waterData','zoo')

lapply(libraries, require, character.only = TRUE)
```

library(dataRetrieval): USGS package that gets streamflow data  direct from the USGS website


<br/>

#### **Step 2**
Data minning.<br>
Try the code below with the site.code here, then use the site code for your watershed.
```{r, warning=FALSE, message=FALSE}
site.code = "04290500"  #  The USGS streamgage code for Winnoski River at Esses
```

<br/>

#### **Step 3**
Download data from the USGS.<br>
```{r, warning=FALSE, message=FALSE}
parameter.code = "00060"  # this is the code for stream discharge.
start.date = "1929-01-01"  # Blanks get all of the data
end.date = "2022-12-31"

# store data in "winnoski dataframe
winooski = readNWISdv("04290500", parameter.code, start.date, end.date)
```


<br/>

#### **Step 4**
Explore date.<br>
```{r, warning=FALSE, message=FALSE}
head(winooski)
tail(winooski)
```


<br/>

#### **Step 5**
The names for the discharge and QA columns aren't very nice, so rename them:<br>
```{r, warning=FALSE, message=FALSE}
names(winooski)[c(4,5)] = c("Q.ft.s","QA")
head(winooski)
```

<br/>

#### **Step 6**
Explore data:<br>
```{r, warning=FALSE, message=FALSE}
plot(winooski$Date,winooski$Q.ft.s, type='l')
```


<br/>

#### **Step 7**
Summarize data and find extreme values.<br>
```{r, warning=FALSE, message=FALSE}
summary(winooski)
winooski[which.max(winooski$Q.ft.s), ] # Identify the maximum flood events in the record
winooski[which.min(winooski$Q.ft.s), ] # Identify the minimum discharge events in the record
```


<br/>

#### **Step 8**
Data visualization.<br>
```{r, warning=FALSE, message=FALSE}
winooski.df <- winooski %>%
  mutate(year = factor(year(Date)),     # use year to define separate curves
         Date = update(Date, year = 1))  # use a constant year for the x-axis


p <-  ggplot(winooski.df, aes(Date, Q.ft.s, color = year)) +
            scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme_bw()

p +  geom_line() + theme(legend.position = "none") # Add the lines. 
```

<br/>

#### **Step 9**
Data visualization.<br>
```{r, warning=FALSE, message=FALSE}
p +  geom_line(aes(group = year), color = "gray20", alpha = 0.1) +
            geom_line(data = function(x) filter(x, year == 1936), color='blue' , size = 1) +
  theme_bw()
```


<br/>

#### **Step 10**
Contrasting 1936 with the rest of the years.<br>
```{r, warning=FALSE, message=FALSE}
year.1936 <- setDT(winooski)[Date %between% c('1936-01-01', '1936-12-31')]
no.2023   <- setDT(winooski)[(Date < '1936-01-01' | Date > '1936-12-31'),]

summary(year.1936)
summary(no.2023)
```


<br/>

#### **Step 11.1**
Calculate cumulative discharge for each year by first grouping by water year.<br>
```{r, warning=FALSE, message=FALSE}
cum.data <- addWaterYear(winooski) # Add a new column with Year
cumulative_dat <- group_by(cum.data, waterYear) %>% # Group by Year
  mutate(cumulative_dis = cumsum(Q.ft.s), # Sum daily discharge (function cumsum())
         wy_doy = seq(1:n())) # Add consecutive day
```

#### **Step 11.1**
First graph.<br>
```{r, warning=FALSE, message=FALSE}
q <- ggplot(cumulative_dat, aes(x = wy_doy, y = cumulative_dis, group = waterYear)) + 
  geom_line(lwd = 0.6) +
  xlab("Julian Day") + ylab("Cumulative dischage (ft.s)") +
  ylim(c(0, 1300000)) +
  xlim(0,366) +
  theme_bw() 
q
```


#### **Step 11.2**
Highlight the rainiest year
```{r, warning=FALSE, message=FALSE}
q + geom_line(data=subset(cumulative_dat, waterYear == "2011"), colour="blue", size=0.9)
```


#### **Step 11.3**
highlight the driest year
```{r, warning=FALSE, message=FALSE}
q + geom_line(data=subset(cumulative_dat, waterYear == "1965"), colour="red", size=0.9) 
```


#### **References**


<br/>
<br/>
