---
title: "Times Series Analysis"
author: 
- "Pablo E. Guti&eacute;rrez-Fonseca"
- "`r format(Sys.time(), '%d %B, %Y')`"
layout: default
permalink:
output: html_document
---
<br/>
<br/>


#### **Step 1**
Load libraries.<br>
```{r, warning=FALSE, message=FALSE}
libraries <- c('Kendall', 'trend', 'changepoint')
lapply(libraries, require, character.only = TRUE)
```

<br/>

#### **Step 2**
```{r, warning=FALSE, message=FALSE}
rain <- read.csv("D:/OneDrive - University of Vermont/Curriculum/19_ Personal webpage/TropicalFreshwaterEcology/_pages/Lectures/data/StreamEcology_TS_precipitation.csv")
rainEVFS <- rain[,3] # Select colunm 3.  
```

<br/>

#### **Step 3**
```{r, warning=FALSE, message=FALSE}
rain.ts <- ts(rainEVFS, # Convert "Precipitation" to a time series object.
                    frequency=12, start=c(1975,1), end=c(2016,12))
head(rain.ts, n=24) 
```

<br/>

#### **Step 4**
```{r, warning=FALSE, message=FALSE}
plot(rain.ts, main="Precipitation El Verde")
```

<br/>

## Mann-Kendall Trend Test
#### **Step 5**
We can use the Mann-Kendall Trend Test to see if there is a pattern in the data.
```{r, warning=FALSE, message=FALSE}
MannKendall(rain.ts)
# tau is a measure of the strength of the trend
# p-value for trend is also shown
```
The test statistic is 0.0301, and the two-sided p-value associated with it is not less than 0.05. We cannot reject the null hypothesis of the test and conclude that no trend exists in the data because the p-value is greater than 0.05.

<br/>

#### **Step 6**
Plot and trend
Now we can add a smooth line to visualize the trend.
```{r, warning=FALSE, message=FALSE}
plot(rain.ts, main="Precipitation El Verde")
lines(lowess(time(rain.ts), rain.ts), col='red')
```

<br/>

## Sen's slope
#### **Step 7**
The sens.slope function in the trend package is used with a time series object.  
The sea.sens.slope function performs the test while taking into account the seasonality of the data.
```{r, warning=FALSE, message=FALSE}
sens.slope(rain.ts)
```

<br/>

## Pettitt's test for change in value 
#### **Step 8**
The pettitt.test function in the trend package identifies a point at which the values in the data change.  The results here suggest that the stage values were higher after May 2009 than before.
```{r, warning=FALSE, message=FALSE}
pettitt.test(rain.ts)
rain[244,]
```

<br/>

## Multiple changes points
#### **Step 9**
The next few commands will be used to find a potential change point in the mean. 
```{r, warning=FALSE, message=FALSE}
mvalueS = cpt.mean(rain.ts, method="BinSeg", Q=5) # BinSeg Binary Segmentation
cpts(mvalueS)
rain[c(7,189,255,420,468),]
plot(mvalueS, cpt.width = 4)
```


<br/>
<br/>
