---
title: "non-Metric Multidimentional Scaling (nMDS)"
author: 
- "Pablo E. Guti&eacute;rrez-Fonseca"
- "`r format(Sys.time(), '%d %B, %Y')`"
layout: default
permalink: /pages/Lectures/Introduction to R/
output: html_document
---
<br/>
<br/>

#### **Step 1**
Load libraries.<br>
_Cargar las librerias que necesitas._
```{r, warning=FALSE,message=FALSE}
libraries <- c("vegan", "ggplot2", "dplyr")
lapply(libraries, require, character.only = TRUE)
```

<br/>

#### **Step 2**
Data<br>
_Cargar los datos._
```{r, results='hide'}
laselva_full  <- read.csv("D:/Curriculum/07_ Cursos/Course_Multivariate_Stat_for_Ecological_Data/data/nMDS_laSelva.csv")
head(laselva_full)

```

<br/>

#### **Step 3** 
_Seleccionar la matrix de especies._
```{r, results='hide'}
laselva_sp <- dplyr::select(laselva_full, Baetid:Collemb)
str(laselva_sp)
ncol(laselva_sp)
nrow(laselva_sp)

```

<br/>

## non-Metric Multidimentional Scaling (nMDS)

#### **Step 1: nMDS**
_Ejecutar un nMDS_
```{r, results = 'hide'}
set.seed(14) # Con este comando, siempre comenzara del mismo lugar.
laselva.mds <- metaMDS(laselva_sp, distance = "bray", k = 2, trymax=100)  #using all the defaults
```

<br>

#### **Step 2: nMDS**
_Resumen de los reusltados_
```{r}
laselva.mds
```
_Lo m&aacute;s importante de ver en este "output" es el **stress**_

As a rule of thumb literature has identified the following cut-ff values for stress-level:<br>
&gt; 0.2 is poor (risks for false interpretation).<br> 
0.1 - 0.2 is fair (some distances can be misleading for interpretation).<br> 
0.05 - 0.1 is good (can be confident in inferences from plot).<br>
&lt; 0.05 is excellent (this can be rare).

<br>

#### **Step 3: nMDS**
Plot the results.
```{r}
plot(laselva.mds, type="t")
```

<br>

#### **Step 4: nMDS**
Stressplot. Large scatter around the line suggests that original dissimilarities are not well
preserved in the reduced number of dimensions. Looks pretty good in this case.
```{r}
stressplot(laselva.mds)
```

<br>

#### **Step 5: nMDS**
_Otra metrica para evaluar el ajuste (buen desempe&ntilde;o) del nMDS es mirar el "Goodness of fit"._
```{r}
gof <- goodness(laselva.mds)
# gof

{plot (laselva.mds, display = 'sites', type = 't', main = 'Goodness of fit') # this function draws NMDS ordination diagram with sites
points (laselva.mds, display = 'sites', cex = 2*gof/mean(gof))} # and this adds the points with size reflecting goodness of fit (bigger = worse fit)
```

<br>

#### **Step 6: nMDS**
_Calcular el % de variacion explicado por cada eje_
```{r}
MDS <- cmdscale(vegdist(laselva_sp, method = "bray"), k = 2, eig = T, add = T )
round(MDS$eig*100/sum(MDS$eig),1)
```

<br>

## Visualizar los datos con ```ggplot2```
<br>

#### **Step 1: Data visualization**
_**Score** son los axis de la ordenacion_
```{r}
data.scores <- as.data.frame(scores(laselva.mds)) 
head(data.scores)
```

<br>

#### **Step 2: Data visualization**
_Agregar las especies al dataframe llamado **data.scores**_
```{r}
species.scores <- as.data.frame(scores(laselva.mds, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
#species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data
```

<br>

#### **Step 3: Data visualization**
_Extraer los meses de la base de datos original (laselva_full) y agregarla a **data.scores**_
```{r}
month  <- dplyr::select(laselva_full, month)
data.scores$month <- unlist(month)
head(data.scores)
```

<br>

#### **Step 4: Data visualization**
_Hay cuatro grupos, previamente identificados por un cluster _
```{r}
grp.a <- data.scores[data.scores$month == "one", ][chull(data.scores[data.scores$month == 
          "one", c("NMDS1", "NMDS2")]),]  # hull values for grp A
grp.b <- data.scores[data.scores$month == "two", ][chull(data.scores[data.scores$month == 
          "two", c("NMDS1", "NMDS2")]),]  # hull values for grp A  
grp.c <- data.scores[data.scores$month == "three", ][chull(data.scores[data.scores$month == 
          "three", c("NMDS1", "NMDS2")]),]  # hull values for grp A  
grp.d <- data.scores[data.scores$month == "four", ][chull(data.scores[data.scores$month == 
          "four", c("NMDS1", "NMDS2")]),]  # hull values for grp A  

hull.data <- rbind(grp.a, grp.b, grp.c, grp.d)
head(hull.data)
```

<br>

#### **Step 5: Data visualization**
```{r}
ggplot() + 
  geom_polygon(data=hull.data,aes(x=NMDS1,y=NMDS2,fill=month,group=month),alpha=0.30) + # add the convex hulls
#  geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2,shape=month,colour=month),size=4) + # add the point markers
#  scale_colour_manual(values=c("one" = "red", "two" = "blue",
#                                "three"="black", "four"="green")) +
  coord_equal() +
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())
```

<br>
<br>
