---
title: "simple RDA"
author: 
- "Pablo E. Gutiérrez-Fonseca"
- "`r format(Sys.time(), '%d %B, %Y')`"
layout: default
output:
  html_document: default
  pdf_document: default
permalink: /pages/Lectures/RDA/
---
<br/>
<br/>

#### **Step 1**
Load libraries.<br>
_Cargar las librerias que necesitas._
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(vegan)
```

<br/>

#### **Step 2**
Data<br>
_Cargar los datos._
```{r}
species=read.csv("D:/Curriculum/07_ Cursos/Course_Multivariate_Stat_for_Ecological_Data/data/RDA_species.csv", header=T, row.names=NULL, sep=",")
env=read.csv("D:/Curriculum/07_ Cursos/Course_Multivariate_Stat_for_Ecological_Data/data/RDA_environmetal_standart.csv", header=T, row.names=NULL, sep=",")
```

<br/>

#### **Step 3**
_Remover la columna de sitos._
```{r}
species_1 <- select(species, -site)
env_1 <- select(env, -site)
```

<br/>

#### **Step 4**
**_Transformar datos._** _Hellinger es una transformaci&aacute;n recomendada por Legendre & Callagher (2001) en datos de abundancia y con una respuesta lineal._
```{r}
species.hel <- decostand(species_1, method = "hellinger")
```

<br/>

### **Seleccion de variables Forward**

```{r}
rda_Forward_Sel <- rda(species.hel~ 1, data= env_1)
rda_var <- rda(species.hel ~ temperature + oxygen + pH +
                conductivity, data= env_1, scale=T)

step_forward <- ordistep(rda_Forward_Sel, scope=formula(rda_var),
                         direction = "forward", perm.max=200, pstep=999)
```

<br/>

## **simple RDA**

#### **Step 1: simple RDA**
_Construir el modelo y ejecutar un RDA simple_
```{r}

rda_tree_all = rda(species.hel ~ temperature + oxygen + pH +
                conductivity, data= env_1, scale=T)

```

<br/>

#### **Step 2: simple RDA**

**Summary**

```{r}

head (summary(rda_tree_all))

```

<br/>

### **Test for colinearity**

Then, we can calculate Variance Inflation Factors (VIF) for each of the constraints (variables) from the environmental matrix. If we find an environmental variable with VIF>10, we'll know that this variable presents colinearity with another or other variables. In that case, we would have to delete the variable from our initial dataset and redo all the analysis. In our example, no variable is redundant with each other (all of them have VIF<10). Linear dependencies can be explored by computing the variables' variance inflation factors (VIF), which measure the proportion by which the variance of a regression coefficient is inflated in the presence of other explanatory variables. VIFs above 20 indicate strong collinearity. Ideally, VIFs above 10 should be at least examined, and avoided if possible. VIFs can be computed in vegan after RDA or CCA (Bocard et al. page 175)

```{r}

vegan::vif.cca(rda_tree_all)

```

<br/>

#### **Step 3: simple RDA**
Plot the results.
```{r}
screeplot(rda_tree_all)
```

<br/>

#### **Step 4: simple RDA**
Percentage explained by constrained and unconstrained variables.
```{r}
constrained_eig <- rda_tree_all$CCA$eig/rda_tree_all$tot.chi*100
unconstrained_eig <- rda_tree_all$CA$eig/rda_tree_all$tot.chi*100
expl_var <- c(constrained_eig, unconstrained_eig)
barplot (expl_var[1:20], col = c(rep ('red', length (constrained_eig)), rep ('black', length (unconstrained_eig))),
         las = 2, ylab = '% variation')

```

<br/>

#### **Step 5: simple RDA**
**Ordination plots**

Scaling 1 (distance biplot): Se prioriza que la distancia entre objetos en el grafico respete tanto como sea posible las distancias euclidias de la matriz original. Los angulos entre vectores (variables) pueden ser distorsionados.

Scaling 2 (correlation biplot): Se prioriza que los angulos entre vectores respeten la correlacion original entre variables. La distancia entre objetos en el grafico puede estar distorsionada.

En resumen, utilizaremos scaling 1 si nos interesa mas ver como se diferencian los objetos, y scaling 2 si nos interesa mas ver como se relacionan las distintas variables.

https://bookdown.org/stephi_gascon/bookdown-demo-master_-_multivariant/_book/ordination.html
https://fukamilab.github.io/BIO202/06-B-constrained-ordination.html

```{r}
par(mfrow=c(1,2))

plot(rda_tree_all, scaling=1, main="Odonata in Urban ponds, scaling=1")
plot(rda_tree_all, scaling=2, main="Odonata in Urban ponds, scaling=2")

```

<br/>

#### **Step 6: simple RDA**
_Calcular las R_

```{r}
(R2 <- vegan::RsquareAdj(rda_tree_all)$r.squared)

(R2adj <- vegan::RsquareAdj(rda_tree_all)$adj.r.squared)

```

<br/>

#### **Step 6: simple RDA**
Testing the global significance of RDA
```{r}
set.seed(1)
vegan::anova.cca(rda_tree_all, step=1000)

```

<br>

Testing the significance of RDA axes (at least the first two or three should present a significant p value)
Hint Argument "step" gives the minimal number of permutations requested 
to assess if the F value of a test is obviously significant or not
```{r}
set.seed(1)
vegan::anova.cca(rda_tree_all, by='axis', step=1000)

```

<br/>

Testing the significance of terms (environmental variables)
Hint Argument "step" gives the minimal number of permutations requested 
to assess if the F value of a test is obviously significant or not
```{r}
vegan::anova.cca(rda_tree_all, by='terms', step=1000)

```

<br>

#### **Glossary**

**Inertia:** is another name for variation or variance in this case.<br>
**"Total":** refers to total variance. Constrained + Unconstrained = Total<br>
**"Constrained":** refers to the amount of variance explained by the explanatory variables.<br>
**"Unconstrained":** refers to the residual variance. 


An R2 statistic can be derived simply as Constrained / Total.

<br>

#### **References**

http://dmcglinn.github.io/quant_methods/lessons/multivariate_models.html

<br/>
<br/>
